CF Registry Analysis

### 2024-05-07 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
tables and plots formatted

## Demographics of Incident CFRD cases versus non-CFRD cases in 2004, 2011, and 2018

```{r}
annual %>%
  filter(ReviewYear %in% c(2004, 2011, 2018)) %>%
  select(
    ReviewYear, CFRD, Age_YrEnd, Gender, Race, Hispanicrace, bmi_perc,
    bmi_perc_cat, MutClass, tube_feeding, liver_disease, A_FEV1_pct_predicted,
    A_FVC_pct_predicted, NumPulmExacerbation, A_corticosteroids1
  ) %>%
  tbl_strata(
    strata = ReviewYear,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(
          by = CFRD, missing_text = "Missing",
          type = list(A_corticosteroids1 ~ "categorical")
        ) %>%
        add_p() %>%
        add_overall() %>%
        separate_p_footnotes()
  )
```

@@

## Kaplan Meier Analysis

```{r}
# Whole cohort
og <- as.numeric(ymd("2003-01-01"))
km_fit <- survfit(
  Surv(as.numeric(tstart), as.numeric(tstop), as.numeric(CFRD) == 2,
    origin = og
  ) ~ 1,
  data = annual, id = eDWID
)
ggsurvplot(km_fit,
  data = annual, risk.table = T, cumevents = T, fun = "event",
  xlab = "Days from 2003-01-01"
)
cum_inc <- tidy(km_fit)
cum_inc$time <- as.Date(cum_inc$time + og)
kable(cum_inc)
# By age group
km_fit <- survfit(
  Surv(as.numeric(tstart), as.numeric(tstop), as.numeric(CFRD) == 2,
    origin = og
  ) ~ age_group,
  data = annual, id = eDWID
)
ggsurvplot(km_fit,
  data = annual, risk.table = T, cumevents = T, fun = "event",
  xlab = "Days from 2003-01-01"
)
km_table <- tidy(km_fit)
km_table$time <- as.Date(km_table$time + og)
km_table <- split.data.frame(km_table, km_table$strata)
```

```{r results='asis'}
invisible(lapply(km_table, function(t) {
  name <- t$strata[1]
  t <- t %>% select(-strata)
  print(kable(t, caption = name))
}))
```

@@

```{r}
annual_split <- survSplit(
  Surv(as.numeric(tstart), as.numeric(tstop), as.numeric(CFRD) == 2,
    origin = og
  ) ~ .,
  data = annual, cut = c(2008, 2013, 2018), episode = "tgroup"
)
annual_split$tgroup <- factor(annual_split$tgroup,
  levels = 1:3, labels = c("[2003,2009)", "[2009,2014)", "[2014,2019)")
)
y <- "Surv(as.numeric(tstart), as.numeric(tstop), event)"
```

>>>>

## Demographics of incident CFRD cases versus non-CFRD cases in 2004, 2011, and 2018

```{r}
#| message: false
annual %>%
  filter(ReviewYear %in% c(2004, 2011, 2018)) %>%
  select(
    ReviewYear, CFRD, Age_YrEnd, Gender, Race, Hispanicrace, bmi_perc,
    bmi_perc_cat, MutClass, tube_feeding, liver_disease, A_FEV1_pct_predicted,
    A_FVC_pct_predicted, NumPulmExacerbation, A_corticosteroids1
  ) %>%
  tbl_strata(
    strata = ReviewYear,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(
          by = CFRD, missing_text = "Missing",
          type = list(A_corticosteroids1 ~ "categorical")
        ) %>%
        add_p() %>%
        add_overall() %>%
        separate_p_footnotes()
  )
```

@@

## Incidence and prevalence

### Full cohort

```{r}
#| message: false
og <- ymd("2003-01-01")
km_fit <- survfit(
  Surv(as.numeric(tstart), as.numeric(tstop), as.numeric(CFRD) == 2,
    origin = as.numeric(og)
  ) ~ 1,
  data = annual, id = eDWID
)
# Plot
surv_p <- ggsurvplot(km_fit, data = annual, fun = "event")
surv_p$plot +
  scale_x_continuous(
    breaks = km_fit$time,
    labels = year(ymd("2003-01-01") + days(km_fit$time))
  ) +
  theme(axis.title.x = element_blank(), legend.position = "none") +
  ggtitle("Cumulative Incidence")
# Table
km_table <- tidy(km_fit)
km_table$time <- year(og + days(km_table$time))
gt(km_table) %>%
  cols_hide("std.error") %>%
  fmt_percent(columns = c("estimate", "conf.low", "conf.high")) %>%
  cols_merge(
    columns = c("estimate", "conf.low", "conf.high"),
    pattern = "{1} [{2} - {3}]"
  ) %>%
  cols_label(
    time = "Year", n.risk = "At Risk", n.event = "CFRD+",
    n.censor = "Censored", estimate = "Estimated CFRD- [95% CI]"
  ) %>%
  opt_row_striping() %>%
  opt_table_lines() %>%
  opt_stylize(style = 6, color = "gray")
```

### By age group

```{r}
#| message: false
# Plot
km_age_fit <- survfit(
  Surv(as.numeric(tstart), as.numeric(tstop), as.numeric(CFRD) == 2,
    origin = as.numeric(og)
  ) ~ age_group,
  data = annual, id = eDWID
)
surv_p <- ggsurvplot(km_age_fit, data = annual, fun = "event")
surv_p$plot +
  scale_x_continuous(
    breaks = km_age_fit$time,
    labels = year(ymd("2003-01-01") + days(km_age_fit$time))
  ) +
  scale_color_discrete(name = "Age Group", labels = c("10-17", "18-29", "30+")) +
  theme(axis.title.x = element_blank(), legend.title = element_blank()) +
  ggtitle("Cumulative Incidence by Age Group")
# Table
km_table <- tidy(km_age_fit)
km_table$time <- year(og + days(km_table$time))
km_table <- split.data.frame(km_table, km_table$strata)

invisible(lapply(km_table, function(km) {
  title <- km$strata[1]
  title <- sub("age_group=", "Ages ", title)
  gt(km) %>%
    cols_hide(c("std.error", "strata")) %>%
    fmt_percent(columns = c("estimate", "conf.low", "conf.high")) %>%
    cols_merge(
      columns = c("estimate", "conf.low", "conf.high"),
      pattern = "{1} [{2} - {3}]"
    ) %>%
    cols_label(
      time = "Year", n.risk = "At Risk", n.event = "CFRD+",
      n.censor = "Censored", estimate = "Estimated CFRD- [95% CI]"
    ) %>%
    opt_row_striping() %>%
    opt_table_lines() %>%
    opt_stylize(style = 6, color = "gray") %>%
    tab_caption(title)
}))
```

## Screening rates

```{r}
screening <- annual %>%
  group_by(ReviewYear, age_group) %>%
  summarise(
    `Screened HbA1c` = mean(!is.na(A_hgba1c)),
    `Screened OGTT` = mean(!is.na(A_twohour_bloodglucose) |
      !is.na(A_ogttfast_bloodglucose)),
    `Screened Either` = mean(!is.na(A_twohour_bloodglucose) |
      !is.na(A_ogttfast_bloodglucose) | !is.na(A_hgba1c)),
    .groups = "drop"
  ) %>%
  pivot_longer(`Screened HbA1c`:`Screened Either`)
ggplot(screening, aes(x = ReviewYear, y = value, color = age_group)) +
  geom_line() +
  facet_wrap(~name) +
  theme_bw() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_color_discrete(name = "Age Group", labels = c("10-17", "18-29", "30+"))
```

@@

```{r}
annual_split <- survSplit(
  Surv(as.numeric(tstart), as.numeric(tstop), as.numeric(CFRD) == 2,
    origin = as.numeric(og)
  ) ~ .,
  data = annual, cut = c(2008, 2013, 2018), episode = "tgroup"
)
annual_split$tgroup <- factor(annual_split$tgroup,
  levels = 1:3, labels = c("[2003,2009)", "[2009,2014)", "[2014,2019)")
)
y <- "Surv(as.numeric(tstart), as.numeric(tstop), event)"
```
